Untitled

### 2024-10-11 analysis/paper/access-ice.qmd
Writes S2+S1 info

All datasets are regrided using bilinear interpolation to a common grid, which is a polar sterographic grid with approximatedly equal area of 25km².

## Description of the model

+S2 is the next saeasonal prediction system after ACCESS-S1.
It uses virtually the same exact model configuration except for some model versions being different.
The main difference are the initial conditions.
While S1 used initial conditions from the UK Met Office, S2 uses the Bureau's own system.
{== Description of the system ==}{\>\>Atmosphere and ocean.
Highlight no ice data assimilation.
Highligh that the hope is that ice will respond to SST/atmosphere DA\<\<}

For evaluation we use hindcst for the period 1981--2023{\>\>Check\<\<}.
Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
Climatology is smoothed with a 11 day running mean.

@@

There is not a lot of data on sea ice properties, especially for things like thickness, age, etc.
However there are relatively reliable satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
These products are not perfect and there still exists a fair bit of observation al uncertainty.
To account for this uncertainty we use multiple datasets.

@@

@@

```{r datasets}
datasets <- list(cdr = CDR(),
                 bt = BT(),
                 era5 = ERA5()
)
which_dataset <- "cdr"
```

@@

## ERA5

```{r dataset_era}
which_dataset <- "era5"
```

```{r}
<<fig-hindcast-extent>>
```

```{r}
<<fig-mean-growth>>
```

```{r}
<<fig-bias>>
```

```{r}
<<<fig-extent-anom>>
```

```{r}
<<fig-rmse>>
```

```{r}
<<fig-iiee>>
```

```{r}
<<fig-iiee-variance>>
```

>>>>

## +S2

+S2 [@wedd2022] is the Australian Bureau of Meteorology's current seasonal prediction system that replaced +S1 [@hudson2017] in October 2021. 
Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0  [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
The atmosphere has a N216 horizontal resolution(~60km in the mid-latitudes) with 85 levels. 
The land model uses the same horizontal grid with 4 soil levels. 
The ocean component has a 1/4º resolution with 75 vertical levels. 
The ice component --based on CICE version 4.1-- has the same resolution than the ocean and 5 sea ice categories as well as an open water category.

Both systems took atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts. 
The main difference between the two are the ocean and ice initial conditions. 
+S1 initial conditions came from from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) DA scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration. 
+S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022. 
This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are larger than 0ºC.
Relevant for this work, sea ice concentrations are not assimilated.

For evaluation we use hindcast for the period 1981--2023{\>\>Check\<\<}.
Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
Climatology is smoothed with a 11 day running mean.

@@

There is not a lot of data on sea ice properties, especially for things like thickness, age, etc.
However there are relatively reliable satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.

We use NOAA/NSIDC's Climate Data Record V4 as the primary verification dataset. 
Its sea ice concentration data are estimated taking the maximum value of the NASA Team [ref] and NASA Bootstrap [ref] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
Both algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites. 
The data has a resolution of 25 by 25km and daily from 1978 onwards.

These data are not perfect, there still exists a fair bit of observational uncertainty.
We include other sources of data to take into acount the observational uncertainty.

However, we find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
Nevertheless, we provide figures using ERA5

The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [ref] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR. 
We also use

Copernicus Climate Change Service (C3S) (2020): Sea ice concentration daily gridded data from 1978 to present derived from satellite observations. Copernicus Climate Change Service (C3S) Climate Data Store (CDS). DOI: 10.24381/cds.3cd8b812 (Accessed on DD-MMM-YYYY)

@@

All datasets are regrided using bilinear interpolation to a common grid, which is a polar stereographic grid with approximately equal area of 25km².

@@

```{r datasets}
datasets <- list(cdr = CDR(),
                 bt = BT(),
                 osi = OSI(),
                 era5 = ERA5()
)
which_dataset <- "cdr"
```

@@

<!-- ## ERA5 -->

<!-- ```{r dataset_era} -->
<!-- which_dataset <- "era5" -->
<!-- ``` -->

<!-- ```{r fig-hindcast-extent-era5} -->
<!-- which_dataset <- "era5" -->
<!-- <<fig-hindcast-extent>> -->
<!-- ``` -->

<!-- ```{r fig-hindcast-extent-bt} -->
<!-- which_dataset <- "bt" -->
<!-- <<fig-hindcast-extent>> -->
<!-- ``` -->

<!-- ```{r fig-mean-growth} -->
<!-- <<fig-mean-growth>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<fig-bias>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<<fig-extent-anom>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<fig-rmse>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<fig-iiee>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<fig-iiee-variance>> -->
<!-- ``` -->
